\speculake: Hiding Malware in Speculative Execution

### 2018-08-07 paper/architecture.tex
pass threat model

We assume a scenario where an adversary wishes to hide or obscure the behavior
of a malicious program (malware) from an analyst attempting to reverse engineer
it. We note this is distinct from the goals of evading malware detection, where
maleware escapes classification by an anti-virus or other automated tool. While
we believe \speculake could also be used to make automated detection more
difficult, our main focus is on reverse engineer resistance, useful for evading
manual classification concerned with malware behavior. For anti-virus evasion,
we refer the reader to several existing techniques that are sufficient to defeat
existing anti-virus
systems~\cite{jana2012abusing,oberheide2009polypack,ropinjector,swinnen2014one,ugarte2015sok}.

@@

We further assume that the malware has a specific trigger that the analyst is
not privy to. In our examples, this trigger is often behavior exhibited by some
other (potentially benign) process running on the same system as the malware.
As the adversary is able to control when such a trigger is deployed (potentially
remotely), the analyst will not be able to observe or force this trigger to
happen at will. We emphasize that while this may also be true for existing
trigger-based malware, analysts can often reverse engineer the trigger out of
the malware, for example by observing control flow within the malware and using
adaptive fuzzers~\cite{afl,stephens2016driller} to generate inputs that explore other execution
paths of the binary. In contrast, \speculake malware's trigger depends on
behavior in seemingly unrelated binaries, influencing malware behavior through
channels that are effectively invisible to the analyst. As with typical malware, the analyst may
attempt to reverse engineer the trigger to reveal the malware's behavior, but we
will show (in Section~\ref{subsec:nested-spec}) how to defeat this type of
analysis.

>>>>

We assume a scenario where an adversary wishes to hide or obscure the behavior
of a malicious program (malware) from an analyst attempting to reverse engineer
it. We note this is distinct from the goals of evading malware detection, where
malware escapes classification by an anti-virus or other automated tool. While
we believe \speculake could also be used to make automated detection more
difficult, our main focus is on reverse engineer resistance, useful for evading
manual classification concerned with malware behavior. For anti-virus evasion,
we refer the reader to several existing techniques that are sufficient to defeat
existing anti-virus
systems~\cite{jana2012abusing,oberheide2009polypack,ropinjector,swinnen2014one,ugarte2015sok}.

@@

We further assume that the malware has a specific trigger that the analyst is
not privy to, and the adversary can influence. In our examples, this
trigger is often behavior exhibited by some other (potentially benign) process
running on the same system as the malware. As the adversary is able to control
when such a trigger is deployed (potentially remotely), the analyst will not be
able to observe or force this trigger to happen at will. We emphasize that while
this may also be true for existing trigger-based malware, analysts can often
reverse engineer the trigger out of the malware, for example by observing
control flow within the malware and using adaptive
fuzzers~\cite{afl,stephens2016driller} to generate inputs that explore other
execution paths of the binary. In contrast, \speculake malware's trigger depends
on behavior in seemingly unrelated binaries, influencing malware behavior
through channels that are effectively invisible to the analyst. As with typical
malware, the analyst may attempt to reverse engineer the trigger to reveal the
malware's behavior, but we will show (in Section~\ref{subsec:nested-spec}) how
to defeat this type of analysis.
